Bayesian exponential random graph modeling of KWL Burial network

### 2020-09-29 analysis/paper/paper.Rmd
making burial map

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures with a focus directly on possible ties between nodes by viewing them as random variables [@Robins2007]. In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measure separately by the presence of a tie without considering interactions among ties [@Morris2008]. By contrast, statistical models assumes social networks as dependent variables and allows a direct modeling for the network dependencies [@Snijders2011]. In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes/ties. One can model network dependencies according to a common experience or research questions to represent relationships between actors through network variables. For example, reciprocity is a variable describing two actors' relationship, and transitivity describes two actors' relationship via a shared third actor. Those predictor variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. The advantage of networks specification under assumptions and contexts is useful to explain archaeological network in a given time. For instance, @Amati2019 demonstrated the use of ERGMs to reconstruct three different network for 15 sites in the Caribbean between AD 100 to 400 to explore the interaction mechanisms. Those models test the hypotheses for the relations between sites in terms of proximity, inter-cultural, and exchange network to explain the presence of hub sites. They found that the observed presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of the hypothesis of proximity or cultural homophily exclusively. However, they also point out some limitations of application, such as static outcome and sensitivity to missing data. In addition, it could be difficult to work on ERGMs because their normalizing constant depending on model parameters is intractable. A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns [@Snijders2006].

>>>>

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical model for expressing properties of network structures with a focus directly on possible ties between nodes by viewing them as random variables [@Robins2007]. In traditional statistical model for network, each predictor variable for forming a tie between two nodes is measure separately by the presence of a tie without considering interactions among ties [@Morris2008]. By contrast, statistical models assumes social networks as dependent variables and allows a direct modeling for the network dependencies [@Snijders2011]. In a network generated by ERGMs, the formation of a tie is influenced by the presence or absence of other ties or attributes of nodes/ties. One can model network dependencies according to a common experience or research questions to represent relationships between actors through network variables. For example, reciprocity is a variable describing two actors' relationship, and transitivity describes two actors' relationship via a shared third actor. Those predictor variables, also called network statistics, are direct functions of the ties themselves where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. The advantage of networks specification under assumptions and contexts is useful to explain archaeological network in a given time. For instance, @Amati2019 demonstrated the use of ERGMs to reconstruct three different network for 15 sites in the Caribbean between AD 100 to 400 to explore the interaction mechanisms. Those models test the hypotheses for the relations between sites in terms of proximity, inter-cultural, and exchange network to explain the presence of hub sites. They found that the observed presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of the hypothesis of proximity or cultural homophily exclusively. However, they also point out some limitations of application, such as static outcome and sensitivity to missing data. In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits ERGMs applications. A Bayesian framework allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].
